---
title: "`r params$report.title`"
---

```{r}
source("impute-mean-function.r")
meth <- meth[which(rownames(meth) %in% sites),]
if (any(is.na(meth))) 
    meth <- impute.mean(meth)
methvar <- rowVars(meth,na.rm=T)
meth <- meth[which(methvar > 2e-16),,drop=F]
```

## Source

```{r, results="asis"}
cat(">", params$dataset.citation, "\n\n>", params$dataset.url, "\n")
```

## Population

```{r}
kableone(CreateTableOne(colnames(vars), data=vars), showAllLevels=T) %>% kable_styling()
```

## Panel CpG sites in the dataset

`r length(intersect(sites, rownames(meth)))`
of the `r length(sites)` 
selected mQTL CpG sites are in the DNAm dataset.

## Clustering by principle components

```{r}
pca.ret <- prcomp(t(meth), scale=T)
```

Associations of ancestry with top 20 principle components
of the DNA methylation dataset (restricted to sites in the panel).

```{r, results="asis"}
stats <- apply(pca.ret$x[,1:20], 2, function(pc) {
    fit <- lm(pc ~ ancestry, data=vars)
    as.data.frame(anova(fit))["ancestry",]
})
stats <- do.call(rbind, stats)
lmable(stats) %>% kable_styling()
```

Plot of three components with the strongest associations

```{r}
best <- order(stats[,ncol(stats)])[1:3]
autoplot(pca.ret, x=best[1], y=best[2], data=vars, colour = 'ancestry')
autoplot(pca.ret, x=best[1], y=best[3], data=vars, colour = 'ancestry')
autoplot(pca.ret, x=best[2], y=best[3], data=vars, colour = 'ancestry')
```

